Liveness-Based Garbage Collection for Lazy Languages

### 2016-04-15 paper/cfg-to-nfa.tex
updated experimental setup

\newcommand{\dfaE}{%
  \scalebox{.7}{
    \psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
    \begin{pspicture}(7,-2)(40,10)
     \putnode{t0}{origin}{5}{0}{} \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \
      }}} \hspace{5mm} \putnode{t2}{t1}{12}{0}{\pscirclebox{\mbox{\  \ \ \
       }}}
       \psset{arrows=->} \ncline{t0}{t1} \ncline{t1}{t2} \putnode{l0}{t1}{5}{2}{\clazy}
         \putnode{tD1}{t2}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\       \       \       \
           }}} 
            \psset{arrows=->} \ncline{t2}{tD1} \putnode{l1}{t2}{5}{2}{\bcar}
             \nccurve[angleA=45,
              angleB=135,                                                 ncurv=4,
              nodesep=-1]{tD1}{tD1} \putnode{lD1}{tD1}{0}{10}{\acdr}
              \end{pspicture} 
              }
              }

\newcommand{\dfaF}{%
  \scalebox{.7}{
    \psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
    \begin{pspicture}(3,-2)(27,10)
     \putnode{t0}{origin}{5}{0}{} \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \
      }}} 
       \hspace{5mm} 
        \putnode{t2}{t1}{12}{0}{\pscirclebox{\mbox{\  \ \ \  }}}
         \psset{arrows=->} \ncline{t0}{t1} \ncline{t1}{t2} \putnode{l0}{t1}{5}{2}{\clazy}
         \end{pspicture}
         }
         }

\newcommand{\dfaG}{%
  \scalebox{.7}{
    \psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
    \begin{pspicture}(3,-2)(16,10)
    \putnode{t0}{origin}{5}{0}{} \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \
      }}} \hspace{1mm} \psset{arrows=->} \ncline{t0}{t1}
      \end{pspicture}
      }
      }

\input{main-pgm1}
For the main  program in Figure~\ref{fig:main-pgm1}, the liveness corresponding
to  the  variable   \var{\Lanv{\pa}{}} at $\pi_{10}$ is given by 
$\var{\Lanv{\pa}{}} \rightarrow \clazy \bcar \var{\Df{\length}{1}}$.
The  DFA   corresponding to this grammar is 
\dfaE. 
As there is no $\acar$ symbol to cancel the $\bcar$ symbol in the automaton no 
$\epsilon$ edges are added and the $\bcar$ edge is deleted to get 
\dfaF.
We finally do the $\clazy$ simplification and notice that there is no path to a final state. 
We delete the $\clazy$ edge without marking any state as final giving the final automata for 
$\var{\Lanv{\pa}{}}$ 
\dfaG.
The final automaton does not
accept any forward paths, reflecting  the lazy nature of our language.
Since
\length\ does not  evaluate the elements of  the argument list,
the  closure for  \pa\ is  never evaluated  and is  reclaimed whenever
liveness-based GC triggers beyond $\pi_9$.

>>>>

\newcommand{\dfaE}{%
  \scalebox{.7}{
    \psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
    \begin{pspicture}(0,-2)(35,8)
     \putnode{t0}{origin}{0}{0}{} \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \
      }}} \hspace{5mm} \putnode{t2}{t1}{12}{0}{\pscirclebox{\mbox{\  \ \ \
       }}}
       \psset{arrows=->} \ncline{t0}{t1} \ncline{t1}{t2} \putnode{l0}{t1}{5}{2}{\clazy}
         \putnode{tD1}{t2}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\       \       \       \
           }}} 
            \psset{arrows=->} \ncline{t2}{tD1} \putnode{l1}{t2}{5}{2}{\bcar}
             \nccurve[angleA=45,
              angleB=135,                                                 ncurv=2,
              nodesep=-1]{tD1}{tD1} \putnode{lD1}{tD1}{0}{10}{\acdr}
              \end{pspicture} 
              }
              }

\newcommand{\dfaF}{%
  \scalebox{.7}{
    \psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
    \begin{pspicture}(3,-2)(24,3)
     \putnode{t0}{origin}{2}{0}{} \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \
      }}} 
       \hspace{5mm} 
        \putnode{t2}{t1}{12}{0}{\pscirclebox{\mbox{\  \ \ \  }}}
         \psset{arrows=->} \ncline{t0}{t1} \ncline{t1}{t2} \putnode{l0}{t1}{5}{2}{\clazy}
         \end{pspicture}
         }
         }

\newcommand{\dfaG}{%
  \scalebox{.7}{
    \psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
    \begin{pspicture}(3,-2)(10,3)
    \putnode{t0}{origin}{2}{0}{} \putnode{t1}{t0}{7}{0}{\pscirclebox{\mbox{\ \ \ \
      }}} \hspace{1mm} \psset{arrows=->} \ncline{t0}{t1}
      \end{pspicture}
      }
      }

\input{main-pgm1}
For the main  program in Figure~\ref{fig:main-pgm1}, the liveness corresponding
to  the  variable   \var{\Lanv{\pa}{}} at $\pi_{10}$ is given by 
$\var{\Lanv{\pa}{}} \rightarrow \clazy \bcar \var{\Df{\length}{1}}$.
The  DFA   corresponding to this grammar is 
\dfaE. 
As there is no $\acar$ symbol to cancel the $\bcar$ symbol in the automaton no 
$\epsilon$ edges are added and the $\bcar$ edge is deleted to get 
\dfaF.
We finally do the $\clazy$ simplification and notice that there is no path to a final state. 
Thus, we delete the $\clazy$ edge without marking any state as final giving the final automata for 
$\var{\Lanv{\pa}{}}$ as 
\dfaG.
The final automaton does not
accept any forward paths, reflecting  the lazy nature of our language.
Since
\length\ does not  evaluate the elements of  the argument list,
the  closure for  \pa\ is  never evaluated  and is  reclaimed whenever
liveness-based GC triggers beyond $\pi_9$.

### 2016-04-15 paper/paper.tex
updated experimental setup

Our experimental setup  consists of an interpreter
for  our  language,  a liveness  analyzer,  and  a
single generation copying  collector.  The garbage
collector can be configured to work on the
basis  of  reachability  (RGC  mode)  or  use  
liveness  DFAs  (LGC  mode).

@@

%\subsection{Garbage collection for references on print stack}
The evaluation of  the top-level expression in a program  is driven by
a printing function~(Section
\ref{sec:semantics}).  
%% Printing of  atomic values is simple  and does
%% not  trigger a  GC.   In case  of  \CONS\ cells,  the
%% \CAR\ is  first printed followed  by \CDR.   In a lazy  language, both
%% \CAR\ and \CDR\  could be closures requiring evaluation,  and this may
%% trigger  more GCs.  When  this happens.   we have  to
%% consider any  references that  might be  on the  print stack  and copy
%% them.
We extend liveness-based  garbage  collection  to
this function.
\section{Experimental evaluation}
\label{sec:experiments}
Our benchmark  consists of programs taken  from nofib~\cite{nofib} and
other repositories for functional programs~\cite{PLT-Scheme, gc_bench,
  huffman-sicp}.     We    ran    the   experiments    on    8    core
Intel\textsuperscript{\textregistered}        Core\textsuperscript{TM}
i7-4770 3.40GHz CPU  having 8192KB L2 cache, 16GB RAM,  running 64 bit
Ubuntu 14.04.

>>>>

@@

%\subsection{Garbage collection for references on print stack}
The evaluation of  the top-level expression in a program  is driven by
a printing function~(Section
\ref{sec:semantics}).  
%% Printing of  atomic values is simple  and does
%% not  trigger a  GC.   In case  of  \CONS\ cells,  the
%% \CAR\ is  first printed followed  by \CDR.   In a lazy  language, both
%% \CAR\ and \CDR\  could be closures requiring evaluation,  and this may
%% trigger  more GCs.  When  this happens.   we have  to
%% consider any  references that  might be  on the  print stack  and copy
%% them.
We extend liveness-based  garbage  collection  to
this function.
\section{Experimental evaluation}
\label{sec:experiments}
Our  experimental  setup   consists  of  the  prototypes   of  (a)  an
interpreter  for our  language, (b)  a  liveness analyzer,  and (c)  a
single generation  copying garbage  collector.  The  garbage collector
can be configured  to work on the basis of  reachability (RGC mode) or
use liveness DFAs (LGC mode).
%
Our benchmark  consists of programs taken  from nofib~\cite{nofib} and
other repositories for functional programs~\cite{PLT-Scheme, gc_bench,
  huffman-sicp}.     We    ran    the   experiments    on    8    core
Intel\textsuperscript{\textregistered}        Core\textsuperscript{TM}
i7-4770 3.40GHz CPU  having 8192KB L2 cache, 16GB RAM,  running 64 bit
Ubuntu 14.04.
